Studies of quantum dots

### 2016-10-23 Manuscript/paper.tex
spEnergiesQD/Manuscript/paper.tex: re-write Wegner generator expression

By expanding the commutator in \eqref{eq:imsrgode} without consideration of the three-body term, we obtain the matrix-element form of the IM-SRG(2) flow equation:
\begin{align}
    \frac{\D H^{[r]}}{\D s} = C^{[r]}(\eta, H) \label{eq:flowmxe}
\end{align}
where
\begin{align}
  C^{[r]}(A, B) &\equiv \tilde{\mathcal{A}}_{A, B} D^{[r]}(A, B) \text{ for any $r$, $A$, $B$} \label{eq:commutmxe} \\
  \tilde{\mathcal{A}}_{x, y} f(x, y) &\equiv f(x, y) - f(y, x) \text{ for any $x$, $y$, $f$}
\end{align}
and finally
\begin{widetext}
  \begin{align}
    D^{[0]}(A, B)
    &\equiv
      \sum_{i, a} n_i \bar{n}_a A^{[1]}_{i, a} B^{[1]}_{a, i}
      + \frac{1}{4} \sum_{i, j, a, b} n_i n_j \bar{n}_a \bar{n}_b A^{[2]}_{i, j, a, b} B^{[2]}_{a, b, i, j}
      \label{eq:flow0} \\
    D^{[1]}_{p, q}(A, B)
    &\equiv
      \sum_{r} A^{[1]}_{p, r} B^{[1]}_{r, q}
      - \frac{1}{2} \sum_{i, j, a} n_i n_j \bar{n}_a A^{[2]}_{i, j, a, q} B^{[2]}_{a, p, i, j}
      + \frac{1}{2} \sum_{i, a, b} n_i \bar{n}_a \bar{n}_b A^{[2]}_{i, p, a, b} B^{[2]}_{a, b, i, q}
      \notag \\
    &\hphantom{=}
      + \sum_{i, a} n_i \bar{n}_a \left(
        A^{[1]}_{i, a} B^{[2]}_{a, p, i, q}
        + A^{[2]}_{i, p, a, q} B^{[1]}_{a, i}
      \right)
      \label{eq:flow1} \\
    D^{[2]}_{p, q, r, s}(A, B)
    &\equiv
      \tilde{\mathcal{A}}_{p, q} \tilde{\mathcal{A}}_{r, s}
        \sum_{i, a} n_i \bar{n}_a A^{[2]}_{i, p, a, r} B^{[2]}_{a, q, i, s}
      + \frac{1}{2} \sum_{i, j} n_i n_j A^{[2]}_{i, j, r, s} B^{[2]}_{p, q, i, j}
      + \frac{1}{2} \sum_{a, b} \bar{n}_a \bar{n}_b A^{[2]}_{p, q, a, b} B^{[2]}_{a, b, r, s}
      \notag \\
    &\hphantom{=}
      + \sum_{t} \left(
        \tilde{\mathcal{A}}_{p, q} A^{[1]}_{q, t} B^{[2]}_{p, t, r, s}
        + \tilde{\mathcal{A}}_{r, s} A^{[2]}_{p, q, r, t} B^{[1]}_{t, s}
      \right)
      \label{eq:flow2}
\end{align}
\end{widetext}
\begin{figure}
\includegraphics[width=.48\textwidth]{fig-diagrams-imsrg}
\caption{Diagrammatic representation of $C^{[r]}(\circ, \bullet)$ in \eqref{eq:commutmxe} that forms part of the IM-SRG flow equation \eqref{eq:flowmxe}.  The diagrams are implicitly antisymmetrized (Hugenholtz diagrams \cite{HUGENHOLTZ1957481} as described in \cite{shavitt2009many} \S 4.4.3).}
\label{fig:diagrams-imsrg}
\end{figure}
The equations can be derived directly from the anticommutation relations of the field operators, but it is more convenient to use Wick's theorem [[ref]] or, even more efficiently, diagrammatic techniques [[ref]] to arrive at the results.  In particular, the diagrammatic representation of antisymmetrized matrix elements are shown in Figure \ref{fig:diagrams-imsrg}.

@@

The original choice suggested by Wegner \cite{PhysRepWegner0} reads
\begin{equation}
  \hat{\eta}_{\text{Wegner}}
  = [\hat{H}^{\text{d}}, \hat{H} - \hat{H}^{\text{d}}]
  = [\hat{H}^{\text{d}}, \hat{H}]
  \label{eq:etaWegner}
\end{equation}
where $\hat{H}^{\text{d}}$ denotes the ``diagonal'' part of the Hamiltonian (and $\hat{H} - \hat{H}^{\text{d}}$ denotes the ``off-diagonal'' part) in the abstract sense described at the end of section \ref{subsubsec:srgmethods}.

The commutator between two Hermitian operators, $\hat{\eta}$ is always antihermitian, as required for a generator.  Additionally, it can be shown that the commutator has the property of suppressing off-diagonal matrix elements as the state evolves via the flow equation \cite{kehrein2006flow}, as we would like.  In particular, matrix elements far off the diagonal, where the Hamiltonian couples states with large energy differences, are suppressed much faster than elements closer to the diagonal.

The matrix elements of the generator can be evaluated using \eqref{eq:commutmxe} to \eqref{eq:flow2}.  If assume that $\hat{H}^{\text{d}}$ does not couple the ground state to the excited states, the equations can be simplified considerably: [[clean up this equation]]
\begin{align*}
\eta_{pq}^{(1)} = & \sum_{r} \left(f_{pr}^d f_{rq} - f_{pr} f_{rq}^d \right) + f_{pq} v_{qppq}^d \left( n_q - n_p \right) \\ \eta_{pqrs}^{(2)} = & - \sum_t \left\lbrace \tilde{\mathcal{A}}_{p, q} f_{pt} v_{tqrs}^d - \tilde{\mathcal{A}}_{r, s} f_{tr} v_{pqts}^d \right\rbrace \notag \\ & + \sum_t \left\lbrace \tilde{\mathcal{A}}_{p, q} f_{pt}^d v_{tqrs} - \tilde{\mathcal{A}}_{r, s} f_{tr}^d v_{pqts} \right\rbrace \notag \\ & + \frac{1}{2} \sum_{tu} (1 - n_t - n_u) \left( v_{pqtu}^d v_{turs} - v_{pqtu} v_{turs}^d \right) \notag \\ & + \sum_{tu} \left( n_t - n_u \right) \tilde{\mathcal{A}}_{p, q} \tilde{\mathcal{A}}_{r, s} v_{tpur}^d v_{uqts},
\end{align*}

@@

The generator is explicitly constructed the following way \cite{PhysRevLett.106.222502,White:cond-mat0201346}
\begin{align}
\hat{\eta}_{\text{White}} &\equiv \hat{\eta}' - \hat{\eta}'{}^\dagger
\end{align}
where
\begin{align}
\eta^{\prime [1]}_{a, i} &\equiv \frac{\bar{n}_a n_i H^{[1]}_{a, i}}{H^{[1]}_{a, a} - H^{[1]}_{i, i} - H^{[2]}_{a, i, a, i}}
\label{eq:WhiteFull1} \\
\eta^{\prime [2]}_{a, b, i, j} &\equiv \frac{4 \bar{n}_a \bar{n}_b n_i n_j H^{[2]}_{a, b, i, j}}{H^{[1]}_{a, a} + H^{[1]}_{b, b} - H^{[1]}_{i, i} - H^{[1]}_{j, j} + X_{a, b, i, j}}
\label{eq:WhiteFull2} \\
X_{a, b, i, j}
  &\equiv v_{a, b, a, b} + v_{i, j, i, j} \notag \\
  &\hphantom{\equiv} - v_{a, i, a, i} - v_{a, j, a, j} \notag \\
  &\hphantom{\equiv} - v_{b, i, b, i} - v_{b, j, b, j}
\label{eq:White7}
\end{align}

>>>>

By expanding the commutator in \eqref{eq:imsrgode} without consideration of the three-body term, we obtain the matrix-element form of the IM-SRG(2) flow equation:
\begin{align}
    \frac{\D H^{[r]}}{\D s} = C^{[r]}(\eta, H) \label{eq:flowmxe}
\end{align}
where
\begin{align}
  C^{[r]}(A, B) &\equiv 2 \mathcal{A}_{A, B} D^{[r]}(A, B) \text{ for any $r$, $A$, $B$} \label{eq:commutmxe} \\
  \mathcal{A}_{x, y} f(x, y) &\equiv \frac{1}{2} \bigl(f(x, y) - f(y, x)\bigr) \text{ for any $x$, $y$, $f$}
\end{align}
and finally
\begin{widetext}
  \begin{align}
    D^{[0]}(A, B)
    &\equiv
      \sum_{i, a} n_i \bar{n}_a A^{[1]}_{i, a} B^{[1]}_{a, i}
      + \frac{1}{4} \sum_{i, j, a, b} n_i n_j \bar{n}_a \bar{n}_b A^{[2]}_{i, j, a, b} B^{[2]}_{a, b, i, j}
      \label{eq:flow0} \\
    D^{[1]}_{p, q}(A, B)
    &\equiv
      \sum_{r} A^{[1]}_{p, r} B^{[1]}_{r, q}
      - \frac{1}{2} \sum_{i, j, a} n_i n_j \bar{n}_a A^{[2]}_{i, j, a, q} B^{[2]}_{a, p, i, j}
      + \frac{1}{2} \sum_{i, a, b} n_i \bar{n}_a \bar{n}_b A^{[2]}_{i, p, a, b} B^{[2]}_{a, b, i, q}
      \notag \\
    &\hphantom{=}
      + \sum_{i, a} n_i \bar{n}_a \left(
        A^{[1]}_{i, a} B^{[2]}_{a, p, i, q}
        + A^{[2]}_{i, p, a, q} B^{[1]}_{a, i}
      \right)
      \label{eq:flow1} \\
    D^{[2]}_{p, q, r, s}(A, B)
    &\equiv
      \tilde{\mathcal{A}}_{p, q} \tilde{\mathcal{A}}_{r, s}
        \sum_{i, a} n_i \bar{n}_a A^{[2]}_{i, p, a, r} B^{[2]}_{a, q, i, s}
      + \frac{1}{2} \sum_{i, j} n_i n_j A^{[2]}_{i, j, r, s} B^{[2]}_{p, q, i, j}
      + \frac{1}{2} \sum_{a, b} \bar{n}_a \bar{n}_b A^{[2]}_{p, q, a, b} B^{[2]}_{a, b, r, s}
      \notag \\
    &\hphantom{=}
      + \sum_{t} \left(
        \tilde{\mathcal{A}}_{p, q} A^{[1]}_{q, t} B^{[2]}_{p, t, r, s}
        + \tilde{\mathcal{A}}_{r, s} A^{[2]}_{p, q, r, t} B^{[1]}_{t, s}
      \right)
      \label{eq:flow2}
\end{align}
\end{widetext}
\begin{figure}
\includegraphics[width=.48\textwidth]{fig-diagrams-imsrg}
\caption{Diagrammatic representation of $C^{[r]}(\circ, \bullet)$ in \eqref{eq:commutmxe} that forms part of the IM-SRG flow equation \eqref{eq:flowmxe}.  The diagrams are implicitly antisymmetrized (Hugenholtz diagrams \cite{HUGENHOLTZ1957481} as described in \cite{shavitt2009many} \S 4.4.3).}
\label{fig:diagrams-imsrg}
\end{figure}
The equations can be derived directly from the anticommutation relations of the field operators, but it is more convenient to use Wick's theorem [[ref]] or, even more efficiently, diagrammatic techniques [[ref]] to arrive at the results.  In particular, the diagrammatic representation of antisymmetrized matrix elements are shown in Figure \ref{fig:diagrams-imsrg}.

@@

The original choice suggested by Wegner \cite{PhysRepWegner0} reads
\begin{equation}
  \hat{\eta}^{\text{Wg}}
  = [\hat{H}^{\text{d}}, \hat{H} - \hat{H}^{\text{d}}]
  = [\hat{H}^{\text{d}}, \hat{H}]
  \label{eq:etaWegner}
\end{equation}
where $\hat{H}^{\text{d}}$ denotes the ``diagonal'' part of the Hamiltonian (and $\hat{H} - \hat{H}^{\text{d}}$ denotes the ``off-diagonal'' part) in the abstract sense described at the end of section \ref{subsubsec:srgmethods}.

Since $\hat{\eta}^{\text{Wg}}$ is the commutator between two Hermitian operators, it is antihermitian as required for a generator.  Additionally, it can be shown that the commutator has the property of suppressing off-diagonal matrix elements as the state evolves via the flow equation \cite{kehrein2006flow}, as we would like.  In particular, matrix elements far off the diagonal, where the Hamiltonian couples states with large energy differences, are suppressed much faster than elements closer to the diagonal.

The matrix elements of the generator can be evaluated using \eqref{eq:commutmxe} to \eqref{eq:flow2}.  If define $\hat{H}^{\text{d}}$ via
\begin{align*}
  H^{\text{d}, [1]}_{p, q} &\equiv \delta_{p, q} H^{[1]}_{p, p} \\
  H^{\text{d}, [2]}_{p, q, r, s} &\equiv (\delta_{p, r} \delta_{q, s} - \delta_{p, s} \delta_{q, r}) H^{[2]}_{p, q, p, q}
\end{align*}
then the expressions for $\hat{\eta}^{\text{Wg}}$ can be simplified considerably:
\begin{align*}
  \eta^{\text{Wg},[0]} &= 0 \\
  \eta^{\text{Wg},[1]}_{p, q} &= 2 H^{[1]}_{p, q} \mathcal{A}_{p, q} \bigl( H^{[1]}_{p, p} + n_p H^{[2]}_{p, q, p, q} \bigr) \\
  \eta^{\text{Wg},[2]}_{p, q, r, s} &= 2 \mathcal{A}_{p, q} \mathcal{A}_{r, s} \Big( \\
  &\qquad H^{[2]}_{p, q, r, s} \mathcal{A}_{(p, q), (r, s)} \Bigl(\\
  &\qquad \qquad 2 H^{[1]}_{p, p} + 4 n_p H^{[2]}_{p, r, p, r} \\
  &\qquad \qquad + (1 - 2 n_p) H^{[2]}_{p, q, p, q} \Bigr) \\
  &\qquad + 4 \delta_{p, r} H^{[1]}_{q, s} \mathcal{A}_{(p, q), (r, s)} H^{[2]}_{p, q, p, q}
  \Bigr)
\end{align*}
where for any $x$, $y$, $z$, $w$, $f$ we define the antisymmetrizer
\begin{align*}
  \mathcal{A}_{(x, y), (z, w)} f(x, y, z, w) &\equiv \frac{1}{2} \bigl(f(x, y, z, w) - f(z, w, x, y)\bigr)
\end{align}

@@

The generator is explicitly constructed the following way \cite{PhysRevLett.106.222502,White:cond-mat0201346}
\begin{align}
\hat{\eta}^{\text{Wh}} &\equiv \hat{\eta}' - \hat{\eta}'{}^\dagger
\end{align}
where
\begin{align}
\eta^{\prime [1]}_{a, i} &\equiv \frac{\bar{n}_a n_i H^{[1]}_{a, i}}{H^{[1]}_{a, a} - H^{[1]}_{i, i} - H^{[2]}_{a, i, a, i}}
\label{eq:WhiteFull1} \\
\eta^{\prime [2]}_{a, b, i, j} &\equiv \frac{4 \bar{n}_a \bar{n}_b n_i n_j H^{[2]}_{a, b, i, j}}{H^{[1]}_{a, a} + H^{[1]}_{b, b} - H^{[1]}_{i, i} - H^{[1]}_{j, j} + X_{a, b, i, j}}
\label{eq:WhiteFull2} \\
X_{a, b, i, j}
  &\equiv H^{[2]}_{a, b, a, b} - H^{[2]}_{a, i, a, i} - H^{[2]}_{b, i, b, i} \notag \\
  &\qquad + H^{[2]}_{i, j, i, j} - H^{[2]}_{a, j, a, j} - H^{[2]}_{b, j, b, j}
\label{eq:White7}
\end{align}
